---
title: Influence of environmental and population factors on Prince William Sound spawning phenology.
author: "Bia Dias"
date: "1/5/2021"
output:
  pdf_document: default
  word_document: default
  html_document:
    df_print: paged
csl: C:/Users/Bia/Dropbox/A_UW/A_PWS_herring_spawningtimming_PAPER/PWS_Herring_phenology/Citations/styles/dependent/marine-ecology-progress-series.csl
always_allow_html: yes
bibliography: C:/Users/Bia/Dropbox/A_UW/A_PWS_herring_spawningtimming_PAPER/PWS_Herring_phenology/Citations/PWS_refs_v1.bib
---
AUTHORS: Dias, B.S., McGowan, D., Campbell, R., Branch, T.

AFFILIATIONS: School of Aquatic and Fishery Sciences, University of Washington

E-MAIL OF CORRESPONDENCE AUTHOR: bdossantosdias@alaska..edu


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(formatR)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60),
                      width=60,tidy=TRUE)

library(tidyverse)
library(mgcv)
library(gamm4)
library(gratia)
library(here)
library(lubridate)
library(lattice) 
library(fitdistrplus)
library(PerformanceAnalytics)
library(olsrr)
library(MASS)
library(AICcmodavg)
library(MuMIn)
library(visreg)
library(kableExtra)
library(equatiomatic)
library(pander)
library(relaimpo)
```


```{r functions, include=FALSE}

Mydotplot <- function(DataSelected){
  
  P <- dotplot(as.matrix(as.matrix(DataSelected)),
               groups=FALSE,
               strip = strip.custom(bg = 'white',
                                    par.strip.text = list(cex = 1.2)),
               scales = list(x = list(relation = "free", draw = TRUE),
                             y = list(relation = "free", draw = FALSE)),
               col=1, cex  = 0.5, pch = 16,
               xlab = list(label = "Value of the variable", cex = 1.5),
               ylab = list(label = "Order of the data from text file", cex = 1.5))
  
  print(P)  
}



Dsquared <- function(model = NULL, 
                     obs = NULL, 
                     pred = NULL, 
                     family = NULL, # needed only when 'model' not provided
                     adjust = FALSE, 
                     npar = NULL) { # needed only when 'model' not provided
  # version 1.4 (31 Aug 2015)
 
  model.provided <- ifelse(is.null(model), FALSE, TRUE)
 
  if (model.provided) {
    if (!("glm" %in% class(model))) stop ("'model' must be of class 'glm'.")
    if (!is.null(pred)) message("Argument 'pred' ignored in favour of 'model'.")
    if (!is.null(obs)) message("Argument 'obs' ignored in favour of 'model'.")
    obs <- model$y
    pred <- model$fitted.values
 
  } else { # if model not provided
    if (is.null(obs) | is.null(pred)) stop ("You must provide either 'obs' and 'pred', or a 'model' object of class 'glm'.")
    if (length(obs) != length(pred)) stop ("'obs' and 'pred' must be of the same length (and in the same order).")
    if (is.null(family)) stop ("With 'obs' and 'pred' arguments (rather than a model object), you must also specify one of two model family options: 'binomial' or 'poisson' (in quotes).")
    else if (!is.character(family)) stop ("Argument 'family' must be provided as character (i.e. in quotes: 'binomial' or 'poisson').")
    else if (length(family) != 1 | !(family %in% c("binomial", "poisson"))) stop ("'family' must be either 'binomial' or 'poisson' (in quotes).")
 
    if (family == "binomial") {
      if (any(!(obs %in% c(0, 1)) | pred < 0 | pred > 1)) stop ("'binomial' family implies that 'obs' data should be binary (with values 0 or 1) and 'pred' data should be bounded between 0 and 1.")
      link <- log(pred / (1 - pred))  # logit
    }  # end if binomial
 
    else if (family == "poisson") {
      if (any(obs %%1 != 0)) stop ("'poisson' family implies that 'obs' data should consist of whole numbers.")
      link <- log(pred)
    }  # end if poisson
 
    model <- glm(obs ~ link, family = family)
  }  # end if model not provided
 
  D2 <- (model$null.deviance - model$deviance) / model$null.deviance
 
  if (adjust) {
    if (model.provided) {
      n <- length(model$y)
      #p <- length(model$coefficients)
      p <- attributes(logLik(model))$df
    } else {
      if (is.null(npar)) stop ("Adjusted D-squared from 'obs' and 'pred' values (rather than a model object) requires specifying the number of parameters in the underlying model ('npar').")
      n <- length(na.omit(obs))
      p <- npar
    }  # end if model.provided else
 
    D2 <- 1 - ((n - 1) / (n - p)) * (1 - D2)
  }  # end if adjust
 
  return (D2)
}

# Guisan, A. & Zimmermann, N.E. (2000) Predictive habitat distribution models in ecology. Ecological Modelling 135: 147-186

```

## Models

### PWS model results #####

```{r fullModel}

PWSdata <- read.csv("DATA_v2/fulldata.csv")

glm_list_full <- list(
glm_full <-  glm(DOY_cum50~ SSB+Mean_age+ ICOADS.SST.apr.wmean + Upwell.wi+
                   NCEP.MerWind.fa+ a.PNA.su, data=PWSdata), 

glm_full_a <-  glm(DOY_cum50~ Mean_age+ ICOADS.SST.apr.wmean + Upwell.wi+
                     NCEP.MerWind.fa+ a.PNA.su, data=PWSdata), 

glm_full_b<-  glm(DOY_cum50~ Mean_age+ ICOADS.SST.apr.wmean + Upwell.wi+ 
                    a.PNA.su, data=PWSdata),
glm_full_c<-  glm(DOY_cum50~ ICOADS.SST.apr.wmean + Upwell.wi+ a.PNA.su, 
                  data=PWSdata),
glm_full_d<-  glm(DOY_cum50~ SSB, data=PWSdata),
glm_full_e<-  glm(DOY_cum50~ Mean_age, data=PWSdata),
glm_full_f<-  glm(DOY_cum50~ ICOADS.SST.apr.wmean, data=PWSdata), 
glm_full_g<-  glm(DOY_cum50~ Upwell.wi, data=PWSdata),
glm_full_h<-  glm(DOY_cum50~ a.PNA.su, data=PWSdata),
glm_full_i <-  glm(DOY_cum50~ Mean_age+ ICOADS.SST.mar.wmean + Upwell.wi+
                     NCEP.MerWind.fa+ a.PNA.su, data=PWSdata)) 
```

```{r fullModel2, echo=FALSE}



fullAICc <- aictab(glm_list_full, modnames = c("~SSB + mean age + April SST + winter downwelling+
                   Fall meridional winds+ Summer PNA", 
                   "~mean age + April SST + winter downwelling+
                   Fall meridional winds+ Summer PNA", 
                   "~mean age + April SST + winter downwelling+Summer PNA",
                   "~April SST + winter downwelling + Summer PNA",
                   "~SSB",
                   "~mean age",
                   "~April SST",
                   "winter downwelling",
                   "Summer PNA", 
                   "~mean age + March SST + winter downwelling+
                   Fall meridional winds+ Summer PNA"
                   ))

################################################
############# Adding Dsquared ################## 

dsqrd_full=NULL
dsAICc_full=NULL

 for(i in 1:length(glm_list_full)){
  dsqrd_full[i]= Dsquared(glm_list_full[[i]])
  dsAICc_full[i]=AICc(glm_list_full[[i]])
 }

dsqrd_full <- as.data.frame(dsqrd_full)
dsAICc_full <- as.data.frame(dsAICc_full)
glmnumbs <- seq_along(1:10)
dsqrd_tab_f <- dsqrd_full %>% 
  mutate(AICc=dsAICc_full$dsAICc_full, ID=glmnumbs) %>% 
  arrange(AICc)


# Create data frame to format into table
tab_01_f = data.frame(fullAICc) %>%
  dplyr::select(
    Modnames, LL, K, AICc, Delta_AICc, AICcWt
  ) %>%
  mutate(
    ER = (max(AICcWt) / AICcWt) %>% round(3) %>% as.character()
  ) %>%
  mutate(
    DE = (dsqrd_tab_f$dsqrd_full) %>% round(3) %>% as.character()
  ) %>% 
   mutate(LL= LL%>% round(3) %>% as.character(),
          AICc= AICc%>% round(3) %>% as.character(),
          AICcWt=AICcWt %>% round(4) %>% as.character(),
          Delta_AICc=Delta_AICc%>% round(3) %>% as.character()) %>% 
  rename(
    Models = Modnames,
    # We can include LaTeX math notation in column names
    # Because \ is a special character we need two \\
    '$\\Delta$AICc' = Delta_AICc,
    'AICc wt.' = AICcWt,
    '$D^{2}$' = DE
  )


pander::pander(tab_01_f)


calc.relimp(glm_full_a, rela= TRUE) # Rela = the sum of the  terms scaled from 0 to 1 

```

### Western model results #####

```{r WesternModelSel}

data_W <- read.csv("DATA_v2/fulldata_W.csv",header = T)

glm_list_W <- list(
glm_W_1<-  glm(DOY_cum50~ SSB, data=data_W), 
glm_W_2<-  glm(DOY_cum50~ MDM_area, data=data_W),
glm_W_3<-  glm(DOY_cum50~ NCEP.SST.fa, 
                  data=data_W),
glm_W_4<-  glm(DOY_cum50~ NCEP.SST.wi, data=data_W),
glm_W_5<-  glm(DOY_cum50~ NCEP.SST.sp, data=data_W), 
glm_W_6<-  glm(DOY_cum50~ Upwell.wi, data=data_W),
glm_W_7<-  glm(DOY_cum50~ a.NPGO.wi, data=data_W),
glm_W_8<-  glm(DOY_cum50~ a.NPGO.su, data=data_W),

glm_W_9<-  glm(DOY_cum50~ SSB + ICOADS.SST.apr.int+
                  NCEP.SST.sp+Upwell.wi+a.NPGO.wi, data=data_W),

glm_W_10<-  glm(DOY_cum50~ SSB + ICOADS.SST.apr.int+
                  Upwell.wi+a.NPGO.su, data=data_W),

glm_W_11<-  glm(DOY_cum50~ SSB + ICOADS.SST.apr.int+
                  Upwell.wi+a.NPGO.wi, data=data_W),


glm_W_12 <- glm(DOY_cum50~ NCEP.SST.sp+ NCEP.SST.wi+ ICOADS.SST.apr.int,
                data=data_W), 


glm_W_13 <- glm(DOY_cum50~ FWinput.total.t+ ICOADS.SST.apr.int,
                data=data_W), 

glm_W_14 <- glm(DOY_cum50~ MDM_area+FWinput.total.t+ ICOADS.SST.apr.int,
                data=data_W), 

glm_W_15 <- glm(DOY_cum50~ SSB+Mean_age+ NCEP.MerWind.wi+
                  FWinput.total.t+ ICOADS.SST.apr.int,
                data=data_W),

glm_W_16 <- glm(DOY_cum50~ Mean_age+
                  FWinput.total.t+ ICOADS.SST.apr.int,
                data=data_W), 

glm_W_17 <- glm(DOY_cum50~ MDM_area+Mean_age+
                  FWinput.total.t+ ICOADS.SST.apr.int,
                data=data_W),

glm_W_18 <- glm(DOY_cum50~ Mean_age+ NCEP.MerWind.wi+
                  FWinput.total.t+ ICOADS.SST.apr.int,
                data=data_W),

glm_W_19 <- glm(DOY_cum50~ Mean_age+ NCEP.MerWind.wi+
                  FWinput.total.t+ ICOADS.SST.apr.int+
                  Upwell.wi+NCEP.MerWind.fa+ a.PNA.su,
                data=data_W),

glm_W_20 <- glm(DOY_cum50~ Mean_age+
                  FWinput.total.t+ ICOADS.SST.apr.int+
                  Upwell.wi+NCEP.MerWind.fa+ a.PNA.su,
                data=data_W),

glm_W_21 <- glm(DOY_cum50~ Mean_age+
                  FWinput.total.t+a.PNA.su,
                data=data_W),## best model

glm_W_22 <- glm(DOY_cum50~ Mean_age+
                  FWinput.total.t+a.MEI.v2.su,
                data=data_W),

glm_W_23 <- glm(DOY_cum50~ Mean_age, data=data_W),

glm_W_24 <- glm(DOY_cum50~ 
                  FWinput.total.t+a.PNA.su,
                data=data_W)

)

```

```{r WesternModelSel2, echo=FALSE}



WAICc <- aictab(glm_list_W, modnames = c(
"~SSB",
"~Mile-days of Milt",
"~Fall offshore SST",
"~Winter offshore SST",
"~Spring offshore SST",
"~winter downwelling",
"~Winter NPGO",
"~Summer NPGO",
"~SSB + April SST + Spring offshore SST + winter downwelling + Winter NPGO",
"~SSB + April SST + winter downwelling+ Summer NPGO",
"~SSB + April SST + winter downwelling + Winter NPGO",
"~Spring offshore SST+ Winter offshore SST + April SST",
"~modeled freshwater input+ April SST",
"~Mile-days of Milt + modeled freshwater input+ April SST",
"~SSB+ mean age + Winter meridional winds + modeled freshwater input + April SST",
"~mean age + modeled freshwater input+ April SST ",
"~Mile-days of Milt + mean age + modeled freshwater input + April SST",
"~mean age + Winter meridional winds + modeled freshwater input + April SST",
"~mean age + Winter meridional winds + modeled freshwater input + April SST + winter downwelling + Fall meridional winds + Summer PNA",
"~mean age + modeled freshwater input + April SST + winter downwelling + Fall meridional winds + Summer PNA",
"~mean age + modeled freshwater input + Summer PNA",
"~mean age + modeled freshwater input + Summer MEI.v2",
"~mean age",
"~modeled freshwater input+Summer PNA"))

################################################
############# Adding Dsquared ################## 

dsqrd_w=NULL
dsAICc_w=NULL

 for(i in 1:length(glm_list_W)){
  dsqrd_w[i]= Dsquared(glm_list_W[[i]])
  dsAICc_w[i]=AICc(glm_list_W[[i]])
 }

dsqrd_w <- as.data.frame(dsqrd_w)
dsAICc_w <- as.data.frame(dsAICc_w)
glmnumbs <- seq_along(1:24)
dsqrd_tab_w <- dsqrd_w %>% 
  mutate(AICc=dsAICc_w$dsAICc_w, ID=glmnumbs) %>% 
  arrange(AICc)

# Create data frame to format into table
tab_02 = data.frame(WAICc) %>%
  dplyr::select(
    Modnames, LL, K, AICc, Delta_AICc, AICcWt
  ) %>%
  mutate(
    ER = (max(AICcWt) / AICcWt) %>% round(3) %>% as.character()
  ) %>%
  mutate(
    DE = (dsqrd_tab_w$dsqrd_w) %>% round(3) %>% as.character()
  ) %>%
  mutate(LL= LL%>% round(3) %>% as.character(),
          AICc= AICc%>% round(3) %>% as.character(),
          AICcWt=AICcWt %>% round(4) %>% as.character(),
          Delta_AICc=Delta_AICc%>% round(3) %>% as.character()) %>%
  rename(
    Models = Modnames,
    # We can include LaTeX math notation in column names
    # Because \ is a special character we need two \\
    '$\\Delta$AICc' = Delta_AICc,
    'AICc wt.' = AICcWt,
    '$D^{2}$' = DE
  )



pander::pander(tab_02)


#diagnostics

calc.relimp(glm_W_21, rela= TRUE) # Rela = the sum of the  terms scaled from 0 to 1 

```

### Eastern model results #####

```{r EasternModelSel}

data_E <- read.csv("DATA_v2/fulldata_E.csv", header = T)

glm_list_E <- list(
glm_E_1<-  glm(DOY_cum50~ SSB, data=data_E), 
glm_E_2<-  glm(DOY_cum50~ MDM_area, data=data_E),  
glm_E_3<-  glm(DOY_cum50~ ICOADS.SST.apr.int, data=data_E), 
glm_E_4<-  glm(DOY_cum50~ NCEP.SST.fa, 
                  data=data_E),
glm_E_5<-  glm(DOY_cum50~ NCEP.SST.wi, data=data_E),
glm_E_6<-  glm(DOY_cum50~ NCEP.SST.sp, data=data_E), 
glm_E_7<-  glm(DOY_cum50~ Upwell.wi, data=data_E),
glm_E_8<-  glm(DOY_cum50~ a.PNA.su, data=data_E),
glm_E_9<-  glm(DOY_cum50~ a.NPGO.su, data=data_E),
glm_E_10<-  glm(DOY_cum50~ NCEP.MerWind.wi, data=data_E),
glm_E_11<-  glm(DOY_cum50~ NCEP.MerWind.fa, data=data_E),
glm_E_12<-  glm(DOY_cum50~  NCEP.ZonWind.wi, data=data_E),
glm_E_13<-  glm(DOY_cum50~  NCEP.ZonWind.sp, data=data_E),

glm_E_14<-  glm(DOY_cum50~ SSB + 
                 NCEP.ZonWind.wi, data=data_E),

glm_E_15<-  glm(DOY_cum50~ SSB + ICOADS.SST.apr.int, data=data_E),

glm_E_16<-  glm(DOY_cum50~ SSB + NCEP.MerWind.wi, data=data_E),

glm_E_17<-  glm(DOY_cum50~ SSB + NCEP.MerWind.fa, data=data_E),

glm_E_18<-  glm(DOY_cum50~ SSB + NCEP.ZonWind.sp, data=data_E),

glm_E_19<-  glm(DOY_cum50~ SSB + ICOADS.SST.apr.int+
                 NCEP.ZonWind.wi, data=data_E),
glm_E_20 <- glm(DOY_cum50~ FWinput.total.t,data=data_E),

glm_E_21 <- glm(DOY_cum50~ SSB + NCEP.MerWind.wi +NCEP.ZonWind.wi+Upwell.wi+
                  a.PNA.su, data=data_E),

glm_E_22 <- glm(DOY_cum50~ MDM_area + NCEP.MerWind.wi +Upwell.wi+
                  a.PNA.su, data=data_E), 

glm_E_23 <- glm(DOY_cum50~ SSB + NCEP.MerWind.wi +Upwell.wi, data=data_E),

glm_E_24 <- glm(DOY_cum50~ Mean_age+NCEP.MerWind.wi +Upwell.wi,
                data=data_E),

glm_E_25 <- glm(DOY_cum50~ Mean_age+
                  FWinput.total.t+ ICOADS.SST.apr.int+
                  Upwell.wi+NCEP.MerWind.fa+ a.PNA.su,
                data=data_E),

glm_E_26 <- glm(DOY_cum50~ MDM_area + Mean_age+
                  NCEP.MerWind.wi+ Upwell.wi, data=data_E),

glm_E_27 <- glm(DOY_cum50~ MDM_area + Mean_age+
                  ICOADS.SST.apr.int+ 
                  NCEP.MerWind.wi+ Upwell.wi+a.PNA.su, data=data_E),

glm_E_28 <- glm(DOY_cum50~ MDM_area + NCEP.MerWind.wi +Upwell.wi+
                  a.PNA.wi, data=data_E),

glm_E_29 <- glm(DOY_cum50~ MDM_area + NCEP.MerWind.fa +Upwell.wi+
                  a.PNA.su, data=data_E)

)

```

```{r EasternModelSel2, echo=FALSE}

EAICc <- aictab(glm_list_E, modnames = c(
"~SSB",
"~Mile-days of Milt",
"~April SST",
"~Fall offshore SST",
"~Winter offshore SST",
"~Spring offshore SST",
"~Winter downwelling",
"~Summer PNA",
"~Summer NPGO",
"~ Winter meridional wind",
"~Fall meridional wind",
"~Winter zonal wind",
"~Spring zonal wind",
"~SSB + Winter zonal wind",
"~SSB + April SST",
"~SSB + Winter meridional wind ",
"~SSB + Fall meridional wind ",
"~SSB + Spring zonal wind ",
"~SSB + April SST + Winter zonal wind",
"~modeled freshwater input",
"~SSB + Winter meridional winds + Winter zonal wind + Winter downwelling + Summer PNA",
"~mile-days of milt + Winter meridional winds + Winter downwelling + Summer PNA",
"~SSB + Winter meridional winds + Winter downwelling",
"~mean age + Winter meridional winds + Winter downwelling",
"~mean age + FWinput.total.t + April SST + Winter downwelling + Fall meridional winds + Summer PNA",
"~mile-days of milt + mean age + Winter meridional winds + Winter downwelling",
"~mile-days of milt + mean age+ICOADS.SST.apr.int+ Winter meridional winds + Winter downwelling + Summer PNA",
"~mile-days of milt + Winter meridional winds + Winter downwelling + Winter PNA",
"~mile-days of milt + Fall meridional winds + Winter downwelling + Summer PNA"))


dsqrd_e=NULL
dsAICc_e=NULL

 for(i in 1:length(glm_list_E)){
  dsqrd_e[i]= Dsquared(glm_list_E[[i]])
  dsAICc_e[i]=AICc(glm_list_E[[i]])
 }

dsqrd_e <- as.data.frame(dsqrd_e)
dsAICc_e <- as.data.frame(dsAICc_e)
glmnumbs <- seq_along(1:29)
dsqrd_tab_e <- dsqrd_e %>% 
  mutate(AICc=dsAICc_e$dsAICc_e, ID=glmnumbs) %>% 
  arrange(AICc)

# Create data frame to format into table
tab_03 = data.frame(EAICc) %>%
  dplyr::select(
    Modnames, LL, K, AICc, Delta_AICc, AICcWt
  ) %>%
  mutate(
    ER = (max(AICcWt) / AICcWt) %>% round(3) %>% as.character()
  ) %>%
  mutate(
    DE = (dsqrd_tab_e$dsqrd_e) %>% round(3) %>% as.character()
  ) %>%
  mutate(LL= LL%>% round(3) %>% as.character(),
          AICc= AICc%>% round(3) %>% as.character(),
          AICcWt=AICcWt %>% round(4) %>% as.character(),
          Delta_AICc=Delta_AICc%>% round(3) %>% as.character()) %>%
  rename(
    Models = Modnames,
    # We can include LaTeX math notation in column names
    # Because \ is a special character we need two \\
    '$\\Delta$AICc' = Delta_AICc,
    'AICc wt.' = AICcWt,
    '$D^{2}$' = DE
  )


pander::pander(tab_03)


#diagnostics
# Relative importance of glm terms- relaimpo::calc.relimp()

calc.relimp(glm_E_22, rela= TRUE) # Rela = the sum of the  terms scaled from 0 to 1 

```


### PWS Western Eastern model results #####


```{r WE_data_model_Sel}

dataPWS <- read_csv(file="DATA_v2/fulldata_PWS.csv")
dataPWS$RS_SSB_10 <- as_factor(dataPWS$RS_SSB_10)
dataPWS$RS_SSB_5 <- as_factor(dataPWS$RS_SSB_5)
dataPWS$RS_SP_10 <- as_factor(dataPWS$RS_SP_10)
dataPWS$RS_SP_5 <- as_factor(dataPWS$RS_SP_5)
dataPWS$Region <- as_factor(dataPWS$Region)

dataPWS_v2 <- dataPWS %>% drop_na() 
dataPWS_v2$Region <- factor(dataPWS_v2$Region, levels = rev(levels(dataPWS_v2$Region)))


#Second interaction of model selection

glm_list1 <- list(
  glm_1_1 <- glm(DOY_cum50~ MDM_area+ NCEP.MerWind.wi +Upwell.wi+
                  a.PNA.su, data=dataPWS_v2),
  
  glm_1_2 <- glm(DOY_cum50~ Mean_age+
                  FWinput.total.t+a.PNA.su,
                data=dataPWS_v2),
  
  glm_1_3 <- glm(DOY_cum50~ MDM_area*Region +Mean_age*Region+ 
                    NCEP.MerWind.wi +Upwell.wi+
                    FWinput.total.t*Region+
                    a.PNA.su,
                data=dataPWS_v2),
  
  glm_1_4 <- glm(DOY_cum50~ Region+MDM_area*Region +Mean_age*Region+ 
                    NCEP.MerWind.wi +Upwell.wi+
                    FWinput.total.t+
                    a.PNA.su,
                data=dataPWS_v2), ##
  
  glm_1_5 <- glm(DOY_cum50~ Region+MDM_area*Region*RS_SSB_10 +Mean_age*Region+ 
                    NCEP.MerWind.wi +Upwell.wi+
                    FWinput.total.t+
                    a.PNA.su,
                data=dataPWS_v2),
  
  glm_1_6 <- glm(DOY_cum50~ Region+ RS_SSB_10+MDM_area*RS_SSB_10+
                    Mean_age*RS_SSB_10+ 
                    NCEP.MerWind.wi +Upwell.wi+
                    FWinput.total.t+
                    a.PNA.su,
                data=dataPWS_v2),
  
  glm_1_7 <- glm(DOY_cum50~ RS_SSB_10+
                    Mean_age*RS_SSB_10+ 
                    NCEP.MerWind.wi +Upwell.wi+
                    FWinput.total.t+
                    a.PNA.su,
                data=dataPWS_v2),
  
  glm_1_8 <- glm(DOY_cum50~ 
                  FWinput.total.t+a.PNA.su,
                data=dataPWS_v2),
  
  glm_1_9 <- glm(DOY_cum50~ 
                  FWinput.total.t*RS_SSB_10,
                data=dataPWS_v2),
  
  glm_1_10 <- glm(DOY_cum50~
                     Mean_age*RS_SSB_10,
                data=dataPWS_v2),
  
  glm_1_11 <- glm(DOY_cum50~ Region+MDM_area*Region + Mean_age*RS_SSB_10+ 
                    NCEP.MerWind.wi +Upwell.wi+
                    FWinput.total.t+
                    a.PNA.su,
                data=dataPWS_v2),

  glm_1_12 <- glm(DOY_cum50~ Region+MDM_area*Region +
                    NCEP.MerWind.wi +Upwell.wi+
                    FWinput.total.t*Region+
                    a.PNA.su,
                data=dataPWS_v2),
  
  glm_WE_13 <- glm(DOY_cum50~ Region+MDM_area*Region +Mean_age*Region+ 
                    NCEP.MerWind.wi + ICOADS.SST.apr.int+ICOADS.SST.mar.intp +Upwell.wi+
                    FWinput.total.t+
                    a.PNA.su,
                data=dataPWS),
  
  glm_WE_14 <- glm(DOY_cum50~ Region+MDM_area*Region +Mean_age*Region+ 
                    NCEP.MerWind.wi + ICOADS.SST.apr.int +Upwell.wi+
                    FWinput.total.t+
                    a.PNA.su,
                data=dataPWS),
  
   glm_WE_15 <- glm(DOY_cum50~ Region+MDM_area*Region +Mean_age*Region+ 
                    NCEP.MerWind.wi + ICOADS.SST.mar.intp +Upwell.wi+
                    FWinput.total.t+
                    a.PNA.su,
                data=dataPWS),
  
  glm_1_0 <- glm(DOY_cum50~1, data=dataPWS_v2),
  glm_1_17 <- glm(DOY_cum50~ Region+MDM_area*Region*RS_SSB_5 +Mean_age*Region+ 
                    NCEP.MerWind.wi +Upwell.wi+
                    FWinput.total.t+
                    a.PNA.su,
                data=dataPWS_v2),
  
  glm_1_18 <- glm(DOY_cum50~ Region+MDM_area*Region*RS_SP_10 +Mean_age*Region+ 
                    NCEP.MerWind.wi +Upwell.wi+
                    FWinput.total.t+
                    a.PNA.su,
                data=dataPWS_v2),
  
  glm_1_19 <- glm(DOY_cum50~ Region+MDM_area*Region*RS_SP_5 +Mean_age*Region+ 
                    NCEP.MerWind.wi +Upwell.wi+
                    FWinput.total.t+
                    a.PNA.su,
                data=dataPWS_v2)
)

```

```{r WE_data_model_Sel2, echo=FALSE}

WEAICc <- aictab(glm_list1, modnames = c(
"~mile-days of milt + Winter meridional wind + Winter downwelling + Summer PNA",
"~mean age + modeled freshwater input + Summer PNA","~mile-days of milt*Region + mean age*region + Winter meridional winds + Winter downwelling + modeled freshwater input*Region+ Summer PNA",
"~region + mile-days of milt *region + mean age*Region + Winter meridional winds + Winter downwelling + modeled freshwater input + Summer PNA",
"~region + mile-days of milt*region*SSB 10-year regime shift +mean age*region + Winter meridional winds + Winter downwelling + modeled freshwater input + Summer PNA",
"~region+ SSB 10-year regime shift + mile-days of milt * SSB 10-year regime shift + mean age* SSB 10-year regime shift + Winter meridional winds + Winter downwelling + modeled freshwater input + Summer PNA",
"~SSB 10-year regime shift + mean age* SSB 10-year regime shift + Winter meridional winds + Winter downwelling + modeled freshwater input + Summer PNA",
"~modeled freshwater input",
"~modeled freshwater input* SSB 10-year regime shift",
"~mean age*SSB 10-year regime shift",
"~region + mile-days of milt *region + mean age * SSB 10-year regime shift + Winter meridional winds + Winter downwelling + modeled freshwater input + Summer PNA",
"~region + mile-days of milt *region + Winter meridional winds + Winter downwelling + modeled freshwater input + Summer PNA",
"~region + mile-days of milt *region + mean age*region+ Winter meridional winds+ April SST + March SST + Winter downwelling + modeled freshwater input + Summer PNA",
"~region + mile-days of milt *region + mean age*region+ Winter meridional winds+ April SST + Winter downwelling + modeled freshwater input + Summer PNA",
"~region + mile-days of milt *region + mean age*region+ Winter meridional winds+ March SST + Winter downwelling + modeled freshwater input + Summer PNA",
"~null model",
"~region+ SSB 5-year regime shift + mile-days of milt * SSB 5-year regime shift + mean age* SSB 5-year regime shift + Winter meridional winds + Winter downwelling + modeled freshwater input + Summer PNA",
"~region+ Surplus production 10-year regime shift + mile-days of milt * Surplus production 10-year regime shift + mean age* Surplus production 10-year regime shift + Winter meridional winds + Winter downwelling + modeled freshwater input + Summer PNA",
"~region+ Surplus production 5-year regime shift + mile-days of milt * Surplus production 5-year regime shift + mean age* Surplus production 5-year regime shift + Winter meridional winds + Winter downwelling + modeled freshwater input + Summer PNA"
))

dsqrd=NULL
dsAICc=NULL

 for(i in 1:length(glm_list1)){
  dsqrd[i]= Dsquared(glm_list1[[i]])
  dsAICc[i]=AICc(glm_list1[[i]])
  print(dsqrd[i])
  print(dsAICc[i])
 }

dsqrd <- as.data.frame(dsqrd)
dsAICc <- as.data.frame(dsAICc)
glmnumbs <- seq_along(1:19)
dsqrd_tab <- dsqrd %>% 
  mutate(AICc=dsAICc$dsAICc, ID=glmnumbs) %>% 
  arrange(AICc)
  
#pander::pander(dsqrd)

# Create data frame to format into table
tab_04 = data.frame(WEAICc) %>%
  dplyr::select(
    Modnames, LL, K, AICc, Delta_AICc, AICcWt
  ) %>%
  mutate(
    ER = (max(AICcWt) / AICcWt) %>% round(3) %>% as.character()
  ) %>% 
  mutate(
    DE = (dsqrd_tab$dsqrd)%>% round(3) %>% as.character()
  ) %>%
  mutate(LL= LL%>% round(3) %>% as.character(),
          AICc= AICc%>% round(3) %>% as.character(),
          AICcWt=AICcWt %>% round(4) %>% as.character(),
          Delta_AICc=Delta_AICc%>% round(3) %>% as.character()) %>%
  rename(
    Models = Modnames,
    # We can include LaTeX math notation in column names
    # Because \ is a special character we need two \\
    '$\\Delta$AICc' = Delta_AICc,
    'AICc wt.' = AICcWt,
    '$D^{2}$' = DE
  )


pander::pander(tab_04)


# Relative importance of glm terms- relaimpo::calc.relimp()

calc.relimp(glm_1_4, rela= TRUE) # Rela = the sum of the  terms scaled from 0 to 1 

```


## Model diagnostics 
 - Collinearity: measured by the variance inflation factor (VIF), which is: 
    $$VIF = \frac{1}{1- R^{2}}$$ 
    
    
```{r model_diag}
# Distribution (Packages MASS and fitdistrplus)
# Cullen and Frey graph
descdist(dataPWS_v2$DOY_cum50, boot=1000)
plotdist(dataPWS_v2$DOY_cum50)
#discrete since we are dealing with day. 
#the distribution is moderately skewed. 

```


```{r model_diag_PWS}

lm_full_a_PWS <-  lm(DOY_cum50~ Mean_age+ ICOADS.SST.apr.wmean + Upwell.wi+
                     NCEP.MerWind.fa+ a.PNA.su, data=PWSdata)
ols_coll_diag(lm_full_a_PWS)

ols_correlations(lm_full_a_PWS)

ols_test_breusch_pagan(lm_full_a_PWS, rhs=T, multiple=T) #heteroscedasticity test


```


```{r model_diag_W}

lm_W_21 <- lm(DOY_cum50~ Mean_age+
                  FWinput.total.t+a.PNA.su,
                data=data_W)


ols_coll_diag(lm_W_21)

ols_correlations(lm_W_21)

ols_test_breusch_pagan(lm_W_21, rhs=T, multiple=T) #heteroscedasticity test


```


```{r model_diag_E}

lm_E_22 <- lm(DOY_cum50~ MDM_area + NCEP.MerWind.wi +Upwell.wi+
                  a.PNA.su, data=data_E)

ols_coll_diag(lm_E_22)

ols_correlations(lm_E_22)

ols_test_breusch_pagan(lm_E_22, rhs=T, multiple=T) #heteroscedasticity test

```






